Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-26 robdml.tex
tex

We propose an additional loss function term to enforce such \emph{intra-class}
structure, as shown in \cref{fig:ics}.
%
Specifically, the anchor $\va$ and its adversarial counterpart $\tilde{\va}$
are separated from the positve sample $\vp$ by reusing the triplet loss,
\ie,
%
\begin{equation}
	L_\text{ICS} = \lambda \cdot L_\text{T}(
	\va, \tilde{\va}, \vp; 0),
\end{equation}
%
where $\lambda$ is a constant weight for this loss term,
and the margin is set as zero to avoid negative effect.
%
The $L_\text{ICS}$ term can be added to the loss function for
adversarial training.

@@

Specifically, we (adversarially) train imagenet-initialized ResNet-18
(RN18)~\cite{resnet} with the output dimension of the last layer changed to
$N{=}512$.
%
The margin $\gamma$ in triplet loss is $0.2$ by default.
%
The setup is also consistent to standard DML~\cite{revisiting}.
%
Adam~\cite{adam} optimizer is used for parameter updates, with a learning rate
as $1.0{\times}10^{-3}$ for $150$ epochs, and a mini-batch size of $112$.
%
Adversarial examples are created under the perturbation budget
$\varepsilon{=}8/255$ in infinity norm, using the
Projected Gradient Descent (PGD)~\cite{madry} method with step size $\alpha{=}1/255$ 
and a default maximum step number $\eta{=}8$.
%

The model performance on benign (\ie, unperturbed) examples is measured in
Recall@1 (R@1), Recall@2 (R@2), mAP and NMI
following~\cite{revisiting,robrank}.
%
The adversarial robustness of a model is measured in Empirical Robustness Score
(ERS)~\cite{robrank}, a normalized score (the higher the better) from
an collection of existing attacks against DML or their simplified white-box
alternatives.
%
Since adversarial training is not gradient masking~\cite{obfuscated}, the
performance of white-box attack can be regarded as the upper bound of the
black-box attacks, and thus a model that is empirically robust to the collection
of white-box attacks is expected to be robust in general.

Note, $u$ is much less than the theoretical upperbound of the loss function,
in order to avoid excessive hardness boost in the functions introduced below.

>>>>

We propose an additional loss function term to enforce such \emph{intra-class}
structure, as shown in \cref{fig:ics}.
%
Specifically, the anchor $\va$ and its adversarial counterpart $\tilde{\va}$
are separated from the positve sample $\vp$ by reusing the triplet loss,
\ie,
%
\begin{equation}
	L_\text{ICS} = \lambda \cdot L_\text{T}(
	\va, \phi(\mA + \hat{\vr}_a), \vp; 0),
\end{equation}
%
where $\lambda$ is a constant weight for this loss term,
and the margin is set as zero to avoid negative effect.
%
The $L_\text{ICS}$ term can be added to the loss function for
adversarial training.

@@

Specifically, we (adversarially) train imagenet-initialized ResNet-18
(RN18)~\cite{resnet} with the output dimension of the last layer changed to
$N{=}512$.
%
The margin $\gamma$ in triplet loss is $0.2$, which is consistent to standard
DML~\cite{revisiting}.
%
Adam~\cite{adam} optimizer is used for parameter updates, with a learning rate
as $1.0{\times}10^{-3}$ for $150$ epochs, and a mini-batch size of $112$.
%
Adversarial examples are created under the perturbation budget
$\varepsilon{=}8/255$ in infinity norm, using PGD~\cite{madry} with step size
$\alpha{=}1/255$ and a default maximum step number $\eta{=}8$.
%
The parameter $u$ is equal to $\gamma$, much less than the upperbound of the
loss function, in order to avoid excessive hardness boost in pseudo-hardness
functions $g_\mathsf{B}$ and $g_\mathsf{LGA}$.

The model performance on benign (\ie, unperturbed) examples is measured in
Recall@1 (R@1), Recall@2 (R@2), mAP and NMI
following~\cite{revisiting,robrank}.
%
The adversarial robustness of a model is measured in Empirical Robustness Score
(ERS)~\cite{robrank}, a normalized score (the higher the better) from
an collection of existing attacks against DML or their simplified white-box
alternatives optimized with PGD ($\eta=32$ for strong effect).
%
Since adversarial training is not ``gradient masking''~\cite{obfuscated}, the
performance of white-box attack can be regarded as the upper bound of the
black-box attacks, and thus a model that is empirically robust to the collection
of white-box attacks is expected to be robust in general.
